\documentclass[preprint]{elsarticle}
\usepackage{changes}
\usepackage{pgf}
\usepackage{comment}
\usepackage[utf8]{inputenc}
\usepackage{graphics,epsfig}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{url}
\usepackage{srcltx}
\usepackage{hyperref}
\usepackage{amsmath,amsfonts ,amssymb,fancyhdr,setspace,lastpage,comment,graphics, cleveref,natbib, lineno}
\usepackage{mathrsfs}
\usepackage{todonotes}
\usepackage[all]{xy}




\DeclareMathOperator{\argmax}{argmax}
\newcommand{\pa}[1]{\left ( #1 \right )}
\newcommand{\abs}[1]{\left | #1 \right |}
\newcommand{\for}[0]{\text{ for }}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
\newcommand{\ip}[2]{\left \langle #1,#2 \right \rangle}
\newcommand{\brak}[1]{\langle #1 \rangle}
\newcommand{\C}[0]{\mathbb{C}}
\newcommand{\Z}[0]{\mathbb{Z}}
\newcommand{\R}[0]{\mathbb{R}}
\newcommand{\norm}[1]{\left \Vert #1 \right \Vert }

\renewcommand{\part}[0]{\partial}

\let\oldphi\phi \let\phi\varphi \let\varphi\oldphi
\let\oldphi\phi \let\epsilon\varepsilon


%\title{Fast adaptive behavior in a tri-trophic system: Breakdown of the paradox of enrichment}
%\title{Solving multispecies population games in continuous space and time}

%\title{Solving multispecies population games in continuous space and time \tnoteref{t1}}\tnotetext[t1]{This document is the results of the researchproject funded by the National Science Foundation.}%\tnotetext[t2]{The second title footnote which is a longertext matter to fill through the whole text width andoverflow into another line in the footnotes area of thefirst page.}


%\author{Emil Friis Frølich, Uffe Høgsbro Thygesen}

\title{Solving multispecies population games in continuous space and time \tnoteref{t1}}
\tnotetext[t1]{This work was supported by the Centre for Ocean Life,
a Villum Kann Rasmussen Centre of Excellence supported
by the Villum Foundation.}
\author[1]{Emil F. Fr{\o}lich\corref{cor1}} %\fnref{fn2}
\ead{jaem@dtu.dk}

\author[2]{Uffe H. Thygesen}
\ead{uhth@dtu.dk}

\cortext[cor1]{Corresponding author}
%\fntext[fn1]{This is the first author footnote.}


\affiliation[1]{organization={Technical University of Denmark, Department of Applied Mathematics and Computer Science - DTU Compute},
     addressline={Building 303B, Matematiktorvet},
     postcode={2800},
     city={Kgs. Lyngby},
     country={Denmark}
     }
\affiliation[2]{organization={Technical University of Denmark, Department of Applied Mathematics and Computer Science - DTU Compute},
      addressline={Building 303B, Matematiktorvet},
      postcode={2800},
      city={Kgs. Lyngby},
      country={Denmark}
      }

\begin{document}



\begin{abstract}
  Game theory has emerged as an important tool to understand interacting populations in the last 50 years. The methods of game theory have been successfully applied to study population dynamics with optimal behavior in simple ecosystem models, but the methods are not   generalizable larger, more complex, systems. Game theory is usually applied population models to determine instantaneous optimal patch distribution, coupling behavior to population dynamics. Complex habitat are usually split into a few patches, but populations in the real world interact in continuous space and time. The assumption of decisions based on perfect information is a large simplification, but is hard to avoid. We develop a game-theoretical method to study population dynamics for interacting populations in continuous space, with instantaneous optimal distribution. A continuous setting allows us to model bounded rationality, and compare the impact of bounded and complete rationality on population dynamics. We use the method study the population dynamics and behavior of forage fish preying on copepods. This interaction daily causes the largest animal migration in the world, the diel vertical migration. Incorporating continuous space and time, we can model the seasonal variation in the migration, separating the effects of light and population numbers. We arrive at  qualitative agreement with empirical findings. Including bounded rationality gives rise to spatial distributions corresponding to reality, while the population dynamics for bounded rationality and complete rationality are equivalent. Our approach is general, and can easily be used for complex terrestrial and aquatic systems.
\end{abstract}
\begin{comment}
\begin{abstract}
  Population dynamics are often modeled without taking behavior into account. This is  in spite of the largest daily feeding times for predators, namely at dawn and dusk, being driven by behavior. The daily pattern stems from the Diel Vertical Migration (DVM) in the aquatic setting and crepuscular behavior in the terrestrial setting. This is usually explained by prey avoiding visual predators, and visual predators seeking to find prey. We develop a game-theoretical model of predator-prey interactions in continuous time and space, finding the Nash equilibrium at every instant. By unifying results for the general resolution of polymatrix games, and a spectral discretization scheme, we can resolve the spatially continuous game nearly instantaneously. Our approach allows a unified model for the slow time-scale of population dynamics, and the fast time-scale of the vertical migration, under seasonal changes. We use the diel vertical migration as a case, examining emergent phenomena from the introduction of the fast dynamics.
  On the behavioral time-scale, we see the emergence of a deep scattering layer from the game dynamics. On the longer time-scale of population dynamics, the introduction of optimal behavior has a strong stabilizing effect, compared to the model without optimal behavior. In a changing seasonal environment, we observe a change in daily migration patterns throughout the seasons, driven by changes in both population and light levels. The framework we propose can easily be adapted to population games in inhomogenous terrestrial environments, and more complex food-webs.
\end{abstract}
\end{comment}
\begin{highlights}\item An easily applicable general model for continuous-time population games in continuous habitats is introduced \item Population dynamics for bounded rationality are indistinguishable from population dynamics with perfect rationality \item Incorporating optimal behavior in a model of population dynamics partially explains seasonal variations in the diel vertical migration \end{highlights}

\begin{keyword} population-dynamics \sep habitat-choice \sep game-theory \sep vertical-migration \sep predator-prey \sep population-game %% keywords here, in the form: keyword \sep keyword
\end{keyword}


\maketitle


%\todo[inline]{Abstracted skal nok skrives om til sidst.}


\linenumbers{}
\modulolinenumbers[3]

\input{introduction_general.tex}
\input{general_model.tex}
\input{specific_example.tex}
\input{results.tex}
\input{discussion.tex}
%\input{conclusion.tex}


\bibliographystyle{elsarticle-harv}
\bibliography{bibliography}

\end{document}
